Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-03-23 ThesisJP/Background/background.tex
Started writing about field simulations. Added hos-field figure.

\section{Volume rendering}\label{sec:volren}
\begin{figure}
\centering
\subfigure[Ray-casting]{
	\includegraphics[width=0.4\textwidth]{img/Volumeraycasting.png}\label{fig:vr_rc}
}
\subfigure[Opacity transfer function]{
	\includegraphics[width=0.5\textwidth]{img/otf.png}\label{fig:vr_otf}
}
\caption{The principle behind volume ray casting. a) A ray is casted out from every pixel in the output image. Along the ray, samples are gathered from the volume. The samples are assigned a color and an opacity before they are combined into one pixel value. b) The opacity transfer function used to determine how opaque a sample should be.}
\label{fig:vr}
\end{figure}
A 3D ultrasound dataset is a cloud of scalar values that somehow needs to be presented to the user. One approach used for cardiac ultrasound  is to extract slices corresponding to different long-axis and short-axis views of the cardiac chambers. Another way is to project the whole volume into a two dimensional image. The latter technique is known as volume rendering, and it present for the user the three dimensional structure of the heart in one two-dimensional image. Many different methods for volume rendering exists \cite{brodlie2001}, and in this thesis we will focus on ray casting \cite{Levoy1988}.

@@

\subsection{Adaptive volume rendering}
As for adaptive beamforming, adaptive volume rendering means to derive and use information about the volume being visualized from the volume it self. In the literature on volume rendering this is known as feature enhancement or visibility driven visualization \cite{viola2005, correa2010visibility, marchesin2010}. The common feature of these techniques is how the sample opacity is modulated in order to achieve visibility of relevant structures. If we somehow are able to classify samples as cardiac tissue we can use this information as input to an adaptive volume rendering algorithm, and by that achieve improved visibility of cardiac tissue.

@@

\section{Ultrasound field simulations}\label{sec:field}
Simulation of ultrasound probes are heavily used by researchers in order to test out new algorithms and probe designs. A  typical simulator for ultrasound probes works by approximating the probe surface with discrete points or patches. In software it is then possible to transmit a pulse from a given probe and measure its response in space and time. The most popular simulator for ultrasound field simulations is Field II \cite{Jensen1992}. All simulations in both \textbf{Paper II} and \textbf{III} are produced with this simulator.

>>>>

\section{Volume rendering}\label{sec:volren}
\begin{figure}[t!]
\centering
\subfigure[Ray-casting]{
	\includegraphics[width=0.4\textwidth]{img/Volumeraycasting.png}\label{fig:vr_rc}
}
\subfigure[Opacity transfer function]{
	\includegraphics[width=0.5\textwidth]{img/otf.png}\label{fig:vr_otf}
}
\caption{The principle behind volume ray casting. a) A ray is casted out from every pixel in the output image. Along the ray, samples are gathered from the volume. The samples are assigned a color and an opacity before they are combined into one pixel value. b) The opacity transfer function used to determine how opaque a sample should be.}
\label{fig:vr}
\end{figure}
A 3D ultrasound dataset is a cloud of scalar values that somehow needs to be presented to the user. One approach used for cardiac ultrasound  is to extract slices corresponding to different long-axis and short-axis views of the cardiac chambers. Another way is to project the whole volume into a two dimensional image. The latter technique is known as volume rendering, and it present for the user the three dimensional structure of the heart in one two-dimensional image. Many different methods for volume rendering exists \cite{brodlie2001}, and in this thesis we will focus on ray casting \cite{Levoy1988}.

@@

\subsection{Adaptive volume rendering}
As for adaptive beamforming, adaptive volume rendering means to derive and use information about the data being visualized from the data itself. In the literature on volume rendering this is known as feature enhancement or visibility driven visualization \cite{viola2005, correa2010visibility, marchesin2010}. The common feature of these techniques is how the sample opacity is modulated in order to achieve visibility of relevant structures. If we somehow are able to classify samples as cardiac tissue we can use this information as input to an adaptive volume rendering algorithm, and by that achieve improved visibility of cardiac tissue.

@@

\section{Ultrasound field simulations}\label{sec:field}
Simulation of ultrasound probes are heavily used by researchers in order to test out new algorithms and probe designs. A simulator for ultrasound probes works by approximating the probe surface with discrete points or patches. In software it is then possible to transmit a pulse from a given probe and measure its response in space and time, the \textit{spatial impulse response}. Figure \ref{fig:hos} shows a snapshot of a simulated pressure field from an unfocused linear array emitting a continuous wave. The image was simulated with the simulator from \textbf{Paper V}. The most popular simulator for medical ultrasound field simulations is Field II \cite{Jensen1992}. All simulations in both \textbf{Paper II} and \textbf{III} are produced with this simulator. The Field II simulator works by dividing each element into mathematical elements, and all field points needs to be in the far-field of these small elements. Then it is possible to approximate the spatial impulse response by summing a trapezoid impulse response function from each mathematical element.  \todo{Bergen flops for Field II. From this we see that calculation of ultrasound simulation is complex.}

\begin{figure}[t!]
\centering
\includegraphics[width=\textwidth]{img/hos.eps}
\caption{An example of a simulated ultrasound pressure field from a linear array without focusing. The field was simulated using the simulator in \textbf{Paper V}.}
\label{fig:hos}
\end{figure}
